Seniority and Hierarchy Configuration Interaction for Radicals and \\ Excited States

### 2023-10-23 Manuscript/Response_Letter/Response_Letter.tex

\alert{
\ce{CN} is the only system for which some of the potential energy curves are not smooth around the equilibrium geometry.
This is mentioned on page x of the revised manuscript, reading
``For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
and somewhat less reliable equilibrium properties in these specific cases.''
}

\clearpage

{4) The authors need to mention (or draw) the molecules of their excited state study. Not everyone might be familiar with the QUEST database. It would be insightful to mention the character of the excited states. I assume these are all singly excited states (no double excitation character?). Can the authors provide additional information regarding their nature (local excitation, charge transfer, Rydberg, etc.)? Specifically, what kind of excited states can be described well or less well with hCI methods?
}

\alert{
The set of systems we considered is now presented in Table II, whreas on page x we describe the character of the excited states, reading
``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.''
The details on each specific excited state can be found in The Supporting Information.
The revised manuscript now mentions the performance of the hCI models for different types of excited state, on page x:
``$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
whereas Rydberg states are better described than valence states (MAEs of \SI{0.14}{\eV} and \SI{0.24}{\eV}).
The same trends are found for the low-order models ($\Delta$hCI1 and $\Delta$hCI1.5), and also for $\Delta$CISD, which slightly outperforms $\Delta$hCI2 for each type of excitation.
The specific MAEs can be found in the Supporting Information.''
This aspect is also discussed for the hCI+EN2 models in the new subsection (IV.E).
}

### 2023-10-23 Manuscript/manuscript.tex
saving work

The various CI models introduced here were further assessed based on calculated vertical excitation energies for 69 electronic states,
from 17 closed-shell systems and 13 radicals \fk{(shown in Table~\ref{tab:systems})}, with geometries extracted from the QUEST database. \cite{Veril_2021}
\fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.}
The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.

\begin{table}[ht!]
\caption{\fk{Systems Considered in the Present Excited State Calculations.}}
\label{tab:systems}
{\color{orange}
\begin{ruledtabular}
\begin{tabular}{lll|lll}
\mc{3}{c|}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
\hline
glyoxal      & \ce{HCF}  & \ce{H2O} & \ce{BeH} & \ce{NH2} & \ce{HCO} \\
acetaldehyde & \ce{HCCl} & \ce{H2S} & \ce{BeF} & \ce{PH2} & \ce{HOC} \\
silylidene   & \ce{HPO}  & \ce{N2}  & \ce{BH2} & vinyl    & \ce{OH}  \\
nitroxyl     & \ce{CF2}  & \ce{NH3} & \ce{CN}  & allyl    & \ce{CO+} \\
ethylene     & \ce{BH}   & \ce{HCl} & \ce{CH3} &          & \\
methanimine  & \ce{BF}   &          &          &          & \\
\end{tabular}
\end{ruledtabular}
}
\end{table}

@@

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  &\fk{CIS+EN2}               & \fk{48} &       &      &      &      \\
CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14  & & & & & & \\
CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08  & & & & & & \\
%
\hline
hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66  & & & & & & \\
hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08  & & & & & & \\
hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07  & & & & & & \\
%
\hline
$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  &\fk{$\Delta$CSF+EN2}       & \fk{50}  &        &      &      &      \\
$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{01}  & \fk{0}.\fk{07} & \fk{0}.\fk{10} & \fk{0}.\fk{10} \\
$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10  & & & & & & \\
$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05  & & & & & & \\
%
\hline
sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
%
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

At the $\Delta$hCI2 level, the unbalance associated with the state-specific classes of determinants is reduced to a great extent.
At this level, the state-specific advantage clearly manifests, with $\Delta$hCI2 having a MAE of \SI{0.20}{\eV}, substantially smaller than the MAE of \SI{3.53}{\eV} obtained from hCI2.
This comparison between ground-state-based and state-specific approaches is analogous to our previous finding on the eCI models of the same order, CISD, and $\Delta$CISD. \cite{Kossoski_2023}
Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD, with MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}.
Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.18}{\eV} against \SI{-0.12}{\eV}).
Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
The slightly worse performance of $\Delta$hCI2 probably stems from a residual unbalance associated with the state-specific classes of determinants.

@@

\fk{[Performance of the hCI models for the different types of transitions.]}

@@

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
}
\label{tab:stat2}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50  & \fk{CIS+EN2}            & \fk{19} &       &      &      &      \\
CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17  & & & & & & \\
CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06  & & & & & & \\
CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08  & & & & & & \\
%
\hline
hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64  & \fk{hCI1+EN2}           & \fk{19} & \fk{+0}.\fk{27} & \fk{0}.\fk{28} & \fk{0}.\fk{36} & \fk{0}.\fk{24} \\
hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68  & \fk{hCI1.5+EN2}         & \fk{19} & \fk{+0}.\fk{22} & \fk{0}.\fk{25} & \fk{0}.\fk{33} & \fk{0}.\fk{24} \\
hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90  & & & & & & \\
hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05  & & & & & & \\
hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05  & & & & & & \\
hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04  & & & & & & \\
%
\hline
$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & \fk{$\Delta$CSF+EN2}    & \fk{19} &       &      &      &      \\
$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & \fk{$\Delta$CISD+EN2}   & \fk{19} & \fk{-0}.\fk{02} & \fk{0}.\fk{04} & \fk{0}.\fk{06} & \fk{0}.\fk{06} \\
$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08  & & & & & & \\
$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44  & \fk{$\Delta$hCI1+EN2}   & \fk{19} & \fk{+0}.\fk{02} & \fk{0}.\fk{17} & \fk{0}.\fk{29} & \fk{0}.\fk{29} \\
$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69  & \fk{$\Delta$hCI1.5+EN2} & \fk{19} & \fk{+0}.\fk{00} & \fk{0}.\fk{10} & \fk{0}.\fk{15} & \fk{0}.\fk{15} \\
$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16  & \fk{$\Delta$hCI2+EN2}   & \fk{19} & \fk{-0}.\fk{00} & \fk{0}.\fk{05} & \fk{0}.\fk{08} & \fk{0}.\fk{08} \\
$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10  & & & & & & \\
$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05  & & & & & & \\
$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02  & & & & & & \\
%
\hline
sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53  & \fk{sCI1+EN2}           & \fk{19} & \fk{+0}.\fk{23} & \fk{0}.\fk{25} & \fk{0}.\fk{32} & \fk{0}.\fk{21} \\
$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37  & \fk{$\Delta$sCI1+EN2}   & \fk{19} & \fk{+0}.\fk{06} & \fk{0}.\fk{12} & \fk{0}.\fk{20} & \fk{0}.\fk{19} \\
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

\fk{[Performance of the sCI models for the different types of transitions.]}

@@

\fk{
We found a similar impact of the EN2 correction for $\Delta$CISD and $\Delta$hCI2.
As discussed above, $\Delta$CISD is somewhat more accurate than $\Delta$hCI2 for the closed-shells (MAEs of \SI{0.17}{\eV} and \SI{0.20}{\eV}),
whereas the EN2 correction decreases their respective MAE to \SI{0.06}{\eV} and \SI{0.07}{\eV}.
In turn, while $\Delta$CISD is slightly less accurate than $\Delta$hCI2 for the open-shells (MAEs of \SI{0.12}{\eV} and \SI{0.10}{\eV}),
the EN2 corrected models show a MAE of \SI{0.04}{\eV} and \SI{0.05}{\eV}, respectively.
Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAE of \SIrange{0.06}{0.07}{\eV}),
and somewhat superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
The same trend is found for $\Delta$CISD+EN2, which slightly outperforms $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above.
}

@@

Finally, we employed different sCI models to compute excitation energies of closed- and open-shell systems.
Despite the four different models employed for the closed-shell excitations, the results are overall disappointing.
For the open-shell transitions, the outcome is more encouraging.
\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than eCI or hCI options.}
The relevance of these sCI models lies in their possible connection with CC methods.
While this is already established for closed-shell systems (pCCD and DOCI deliver very similar energies), \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967,Kossoski_2021,Marie_2021}
here we raised the question of whether there exists a polynomial scaling extension of pCCD to open-shell systems that matches sCI1 in terms of computed energies.
In this sense, the present results obtained with $\Delta$sCI1 for open-shell excitations are appealing and encourage the development of a generalized pCCD method.

>>>>

The various CI models introduced here were further assessed based on calculated vertical excitation energies for 69 electronic states,
from 17 closed-shell systems and 13 radicals \fk{(shown in Table~\ref{tab:systems})}, with geometries extracted from the QUEST database. \cite{Veril_2021}
\fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.}
The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.

\begin{table}[ht!]
\caption{\fk{Systems Considered in the Excited State Calculations.}}
\label{tab:systems}
{\color{orange}
\begin{ruledtabular}
\begin{tabular}{lll|lll}
\mc{3}{c|}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
\hline
glyoxal      & \ce{HCF}  & \ce{H2O} & \ce{BeH} & \ce{NH2} & \ce{HCO} \\
acetaldehyde & \ce{HCCl} & \ce{H2S} & \ce{BeF} & \ce{PH2} & \ce{HOC} \\
silylidene   & \ce{HPO}  & \ce{N2}  & \ce{BH2} & vinyl    & \ce{OH}  \\
nitroxyl     & \ce{CF2}  & \ce{NH3} & \ce{CN}  & allyl    & \ce{CO+} \\
ethylene     & \ce{BH}   & \ce{HCl} & \ce{CH3} &          & \\
methanimine  & \ce{BF}   &          &          &          & \\
\end{tabular}
\end{ruledtabular}
}
\end{table}

@@

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  & & & & & & \\
CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14  & & & & & & \\
CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08  & & & & & & \\
%
\hline
hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66  & & & & & & \\
hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08  & & & & & & \\
hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07  & & & & & & \\
%
\hline
$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  & & & & & & \\
$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{01}  & \fk{0}.\fk{07} & \fk{0}.\fk{10} & \fk{0}.\fk{10} \\
$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10  & & & & & & \\
$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05  & & & & & & \\
%
\hline
sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
%
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

At the $\Delta$hCI2 level, the unbalance associated with the state-specific classes of determinants is reduced to a great extent.
At this level, the state-specific advantage clearly manifests, with $\Delta$hCI2 having a MAE of \SI{0.20}{\eV}, substantially smaller than the MAE of \SI{3.53}{\eV} obtained from hCI2.
This comparison between ground-state-based and state-specific approaches is analogous to our previous finding on the eCI models of the same order, CISD, and $\Delta$CISD. \cite{Kossoski_2023}
Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD, with MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}.
Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.18}{\eV} against \SI{-0.12}{\eV}).
Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
The slightly worse performance of $\Delta$hCI2 probably stems from a residual unbalance associated with the state-specific classes of determinants.
\fk{$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}), 
whereas Rydberg states are better described than valence states (MAEs of \SI{0.14}{\eV} and \SI{0.24}{\eV}).
The same trends are found for the low-order models ($\Delta$hCI1 and $\Delta$hCI1.5), and also for $\Delta$CISD, which slightly outperforms $\Delta$hCI2 for each type of excitation.
The specific MAEs can be found in the {\SupInf}.}

@@

@@

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
}
\label{tab:stat2}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50  & & & & & & \\
CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17  & & & & & & \\
CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06  & & & & & & \\
CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08  & & & & & & \\
%
\hline
hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64  & \fk{hCI1+EN2}           & \fk{19} & \fk{+0}.\fk{27} & \fk{0}.\fk{28} & \fk{0}.\fk{36} & \fk{0}.\fk{24} \\
hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68  & \fk{hCI1.5+EN2}         & \fk{19} & \fk{+0}.\fk{22} & \fk{0}.\fk{25} & \fk{0}.\fk{33} & \fk{0}.\fk{24} \\
hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90  & & & & & & \\
hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05  & & & & & & \\
hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05  & & & & & & \\
hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04  & & & & & & \\
%
\hline
$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & & & & & & \\
$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & \fk{$\Delta$CISD+EN2}   & \fk{19} & \fk{-0}.\fk{02} & \fk{0}.\fk{04} & \fk{0}.\fk{06} & \fk{0}.\fk{06} \\
$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08  & & & & & & \\
$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44  & \fk{$\Delta$hCI1+EN2}   & \fk{19} & \fk{+0}.\fk{02} & \fk{0}.\fk{17} & \fk{0}.\fk{29} & \fk{0}.\fk{29} \\
$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69  & \fk{$\Delta$hCI1.5+EN2} & \fk{19} & \fk{+0}.\fk{00} & \fk{0}.\fk{10} & \fk{0}.\fk{15} & \fk{0}.\fk{15} \\
$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16  & \fk{$\Delta$hCI2+EN2}   & \fk{19} & \fk{-0}.\fk{00} & \fk{0}.\fk{05} & \fk{0}.\fk{08} & \fk{0}.\fk{08} \\
$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10  & & & & & & \\
$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05  & & & & & & \\
$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02  & & & & & & \\
%
\hline
sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53  & \fk{sCI1+EN2}           & \fk{19} & \fk{+0}.\fk{23} & \fk{0}.\fk{25} & \fk{0}.\fk{32} & \fk{0}.\fk{21} \\
$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37  & \fk{$\Delta$sCI1+EN2}   & \fk{19} & \fk{+0}.\fk{06} & \fk{0}.\fk{12} & \fk{0}.\fk{20} & \fk{0}.\fk{19} \\
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

@@

\fk{
We found a similar impact of the EN2 correction for $\Delta$CISD and $\Delta$hCI2.
As discussed above, $\Delta$CISD is somewhat more accurate than $\Delta$hCI2 for the closed-shells (MAEs of \SI{0.17}{\eV} and \SI{0.20}{\eV}),
whereas the EN2 correction decreases their respective MAE to \SI{0.06}{\eV} and \SI{0.07}{\eV}.
In turn, while $\Delta$CISD is slightly less accurate than $\Delta$hCI2 for the open-shells (MAEs of \SI{0.12}{\eV} and \SI{0.10}{\eV}),
the EN2 corrected models show a MAE of \SI{0.04}{\eV} and \SI{0.05}{\eV}, respectively.
Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAEs of \SI{0.07}{\eV} and \SI{0.06}{\eV}),
and superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
$\Delta$CISD+EN2 displays the same trends, slightly outperforming $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above, just as discussed above for the unperturbed case.
}

@@

Finally, we employed different sCI models to compute excitation energies of closed- and open-shell systems.
Despite the four different models employed for the closed-shell excitations, the results are overall disappointing.
For the open-shell transitions, the outcome is more encouraging.
\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than the eCI or hCI options.}
The relevance of these sCI models lies in their possible connection with CC methods.
While this is already established for closed-shell systems (pCCD and DOCI deliver very similar energies), \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967,Kossoski_2021,Marie_2021}
here we raised the question of whether there exists a polynomial scaling extension of pCCD to open-shell systems that matches sCI1 in terms of computed energies.
In this sense, the present results obtained with $\Delta$sCI1 for open-shell excitations are appealing and encourage the development of a generalized pCCD method.
